---
title: "Benchmark"
output:
  html_document:
    code_folding: hide
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The output of `versus::compare()` is designed to contain only the minimal amount of information needed to represent the difference between two data frames: rows with value differences, unmatched columns, and unmatched rows. While other packages offer much more functionality, the minimal approach makes versus much faster.

Below is an example benchmark using data from the `nycflights13` package over-sampled to various sizes. In the example data, 5% of rows from each data frame are missing in the other, and in 4 out of the 15 columns, 5% of values are different.

At 1 million rows, versus takes 0.6 seconds and less than 1GB of memory while others take over 10 seconds and multiple GB of memory.

### Benchmark code:

```{r, warning=FALSE, message=FALSE, eval=FALSE}
suppressPackageStartupMessages(library(dplyr))

generate_example_data <- function(n_rows) {
  tbl <- nycflights13::weather |>
    slice_sample(n = n_rows, replace = TRUE) |>
    # simulate unique key
    mutate(time_hour = Sys.time() + row_number())

  tbl_a <- tbl |>
    slice_sample(n = floor(n_rows * 0.95))
  tbl_b <- tbl |>
    slice_sample(n = floor(n_rows * 0.95)) |>
    # simulate value differences
    mutate(across(temp:wind_dir, \(x) ifelse(runif(n()) < .05, x + 1, x)))
  list(a = tbl_a, b = tbl_b)
}

bench_out <- bench::press(
  n_rows = c(1e5, 1e6, 2e6),
  {
    tbl <- generate_example_data(n_rows)
    bench::mark(
      versus =
        versus::compare(tbl$a, tbl$b, by = c(origin, time_hour)),
      arsenal =
        arsenal::comparedf(tbl$a, tbl$b, by = c("origin", "time_hour")),
      dataCompareR =
        dataCompareR::rCompare(tbl$a, tbl$b, keys = c("origin", "time_hour")),
      diffdf =
        diffdf::diffdf(tbl$a, tbl$b, keys = c("origin", "time_hour"), suppress_warnings = TRUE),
      min_iterations = 2,
      max_iterations = 2,
      check = FALSE
    )
  }
)
```

### Benchmark results:

```{r, eval=FALSE}
# benchmark run on 2020 i7 MBP with 32GB of memory
bench_out |>
  summary() |>
  mutate(across(n_rows, scales::comma)) |>
  select(1:6) |>
  print(n = Inf)
#> # A tibble: 12 × 6
#>    expression   n_rows         min   median `itr/sec` mem_alloc
#>    <bch:expr>   <chr>     <bch:tm> <bch:tm>     <dbl> <bch:byt>
#>  1 versus       100,000    59.49ms  62.94ms  15.9       64.58MB
#>  2 arsenal      100,000   958.93ms    1.09s   0.913    388.97MB
#>  3 dataCompareR 100,000      1.41s    1.43s   0.698    265.37MB
#>  4 diffdf       100,000      4.73s    4.75s   0.211    554.93MB
#>  5 versus       1,000,000 536.28ms 586.81ms   1.70     648.99MB
#>  6 arsenal      1,000,000   12.86s   13.08s   0.0765     3.89GB
#>  7 dataCompareR 1,000,000   17.16s   17.25s   0.0580     2.63GB
#>  8 diffdf       1,000,000    56.2s   56.93s   0.0176     5.35GB
#>  9 versus       2,000,000    1.35s    1.37s   0.732      1.27GB
#> 10 arsenal      2,000,000   29.47s    30.5s   0.0328     7.54GB
#> 11 dataCompareR 2,000,000   36.16s   37.32s   0.0268     5.27GB
#> 12 diffdf       2,000,000    2.03m    2.09m   0.00796   10.83GB
```
