---
title: "Benchmark"
output:
  html_document:
    code_folding: hide
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The output of `versus::compare()` is designed to contain only the minimal amount of information needed to represent the difference between two data frames: value differences, unmatched columns, and unmatched rows. While other packages offer much more functionality, the minimal approach makes versus much faster.

Below is an example benchmark using data from the `nycflights13` package over-sampled to various sizes. In the example data, 5% of rows from each data frame are missing in the other, and in 4 out of the 15 columns, 5% of values are different.

At 1 million rows, versus takes 0.6 seconds and less than 1GB of memory while others take over 10 seconds and multiple GB of memory.

### Benchmark code:

```{r, warning=FALSE, message=FALSE, eval=FALSE}
suppressPackageStartupMessages(library(dplyr))

generate_example_data <- function(n_rows) {
  tbl <- nycflights13::weather |>
    slice_sample(n = n_rows, replace = TRUE) |>
    # simulate unique key
    mutate(time_hour = Sys.time() + row_number())

  tbl_a <- tbl |>
    slice_sample(n = n_rows * 0.95)
  tbl_b <- tbl |>
    slice_sample(n = n_rows * 0.95) |>
    # simulate value differences
    mutate(across(temp:wind_dir, \(x) ifelse(runif(n()) < .05, x + 1, x)))
  list(a = tbl_a, b = tbl_b)
}

bench_out <- bench::press(
  n_rows = c(1e5, 1e6, 2e6),
  {
    tbl <- generate_example_data(n_rows)
    bench::mark(
      versus =
        versus::compare(tbl$a, tbl$b, by = c(origin, time_hour)),
      arsenal =
        arsenal::comparedf(tbl$a, tbl$b, by = c("origin", "time_hour")),
      dataCompareR =
        dataCompareR::rCompare(tbl$a, tbl$b, keys = c("origin", "time_hour")),
      diffdf =
        diffdf::diffdf(tbl$a, tbl$b, keys = c("origin", "time_hour"), suppress_warnings = TRUE),
      min_iterations = 2,
      max_iterations = 2,
      check = FALSE
    )
  }
)
```

### Benchmark results:

```{r, eval=FALSE}
# benchmark run on 2020 i7 MBP with 32GB of memory
bench_out |>
  summary() |>
  mutate(across(n_rows, scales::comma)) |>
  select(1:6) |>
  print(n = Inf)
#> # A tibble: 12 × 6
#>    expression   n_rows         min   median `itr/sec` mem_alloc
#>    <bch:expr>   <chr>     <bch:tm> <bch:tm>     <dbl> <bch:byt>
#>  1 versus       100,000    69.66ms  70.23ms  14.2       65.03MB
#>  2 arsenal      100,000   748.77ms 760.06ms   1.32     389.01MB
#>  3 dataCompareR 100,000      1.16s    1.22s   0.823    265.43MB
#>  4 diffdf       100,000      3.87s     4.3s   0.232    554.98MB
#>  5 versus       1,000,000 533.15ms 591.89ms   1.69     653.39MB
#>  6 arsenal      1,000,000   12.51s   12.58s   0.0795     3.77GB
#>  7 dataCompareR 1,000,000   15.26s   15.98s   0.0626     2.63GB
#>  8 diffdf       1,000,000   52.24s   55.23s   0.0181     5.35GB
#>  9 versus       2,000,000    1.23s    1.26s   0.795      1.27GB
#> 10 arsenal      2,000,000   26.23s   26.79s   0.0373     7.54GB
#> 11 dataCompareR 2,000,000   35.36s   35.42s   0.0282     5.27GB
#> 12 diffdf       2,000,000    1.87m    1.96m   0.00849    10.7GB
```
